Analysing Melbourne CBD pedestrian counts data with a focus on imputation methods

### 2017-10-22 thesis_draft/01-chap1.Rmd
updated intro

Imputation methods are explored due to the lack of complete data in the pedestrian dataset.

@@

Throughout the data exploration and analysis, issues with respect to missing values in the pedestrian data are explored and solutions to these issues are proposed. These issues have been seen to cause biased estimates when building models for prediction. Efficient methods for solving these issues is also a priority due to the size of the data.

@@

The dataset being analysed can be obtained from the official City of Melbourne's open data[@com_data]. The pedestrian data is in the form of hourly pedestrian counts for `r length(unique(ped_data$Sensor_ID))` sensor locations. In total, there is `r nrow(ped_data)` observations in the dataset, representing `r nrow(dfa)` hours of data. It is available in `.csv` format, allowing for easy data import into R. We can also access the pedestrian data using the the `rwalkr` package [@_wang_rwalkr]. This package allows R to import the data from `data.melbourne.vic.gov.au` which is updated monthly, or from the data source which is used by `pedestrian.melbourne.vic.gov.au` which is updated daily.

@@

@@

### Literature Review
<!-- #### IMPUTATION OF MISSING CLASSIFIED TRAFFIC DATA DURING WINTER SEASON - _Hyuk-Jae Roh, Satish Sharma, Prasanta K. Sahu_   -->

<!-- http://ir.lib.uwo.ca/cgi/viewcontent.cgi?article=1250&context=csce2016   -->
<!-- - traffic counters with 40% to 60% missing values -->
<!-- - heuristic methods -->
<!--   - using "good historical values" or historical average values resulted -->
<!--     in MARE (mean absolute relative error) reaching up to 80% -->
<!--   - using moving average values worked better than other heuristic methods -->
<!--   - all heuristic methods inhenrently cannot reflect sudden fluctuations during abnormal periods -->
<!-- - pattern matching methods -->
<!--   - find "candidate" volume pattern to compare to "study curve" and use the candidate which best matches the study curve -->
<!-- - ARIMA -->
<!--   - did not work well with multiple seasonalities -->
<!-- - AI methods -->
<!--   - genetic algorithms (GAs) and artificial neural networks (ANNs) -->

<!-- Method used in the study was kNN with an interaction using a weather model, which they found to work best with large variations in traffic during winter -->
<!-- (severe weather conditions affecting traffic volume)   -->

@@

For the purposes of that paper, it was found to be inappropriate with the data being analysed.  However, unlike the highway traffic data, where patterns were compared to traffic volumes in different jurisdictions (large geographical distance), the geographic distances between the sensor locations in the Melbourne CBD pedestrian data are much smaller. Another major difference is the randomness of the missing values, where the periods of missing data run longer in the Melbourne CBD pedestrian data compared to that in the Alberta traffic data. Roh et al. proposed using non-parametric estimation methods with k-Nearest Neighbours (kNN) for imputation, although this is unlikely to work well with the lack of data at certain sensor locations in the Melbourne CBD Pedestrian data.

>>>>

Imputation methods are explored due to the lack of complete data in the pedestrian dataset. Without imputation on the data, some sensors lack information about pedestrian activity. This makes statistical inference and model estimation difficult to compare between locations.

@@

Throughout the data exploration and analysis, issues with respect to missing values in the pedestrian data are explored and solutions to these issues are proposed. These issues have been seen to cause biased estimates when building models for prediction. Efficient methods for solving these issues is also a priority due to the size of the data. As such, the imputation method used should be computationally feasible.

@@

The dataset being analysed can be obtained from the official City of Melbourne's online open data platform. The pedestrian data is in the form of hourly pedestrian counts for `r length(unique(ped_data$Sensor_ID))` sensor locations. In total, there is `r nrow(ped_data)` observations in the dataset, representing `r nrow(dfa)` hours of data. It is available in `.csv` format, allowing for easy data import into R. Access ti the pedestrian data using the the `rwalkr` package. This package allows R to import the data from `data.melbourne.vic.gov.au` which is updated monthly, or from the data source which is used by `pedestrian.melbourne.vic.gov.au` which is updated daily.

@@

The data source used in this paper is accessed with the `rwalkr::walk_melb()` function, which imports data from `pedestrian.melbourne.vic.gov.au`. This data source was chosen due to missing values being explicit (all possible rows in the data are given, even if data is missing). It is also updated more frequently.

@@

The sensor data is collected by sensors located throughout the Melbourne CBD under awnings and street poles. The sensors record the number of people passing through a counting zone each hour, sending this data to a server to be saved. The date of installation varies, as new sensors are constantly being installed over time.

### Literature Review

@@

For the purposes of that paper, it was found to be inappropriate with the data being analysed.  However, unlike the highway traffic data, where patterns were compared to traffic volumes in different jurisdictions (large geographical distance), the geographic distances between the sensor locations in the Melbourne CBD pedestrian data are much smaller. Another major difference is the randomness of the missing values, where the periods of missing data run longer in the Melbourne CBD pedestrian data compared to that in the Alberta traffic data. Roh et al. proposed using non-parametric estimation methods with k-Nearest Neighbours (kNN) for imputation, although this is unlikely to work well with the lack of data at certain sensor locations in the Melbourne CBD Pedestrian data. Similarly, most other applications of imputation methods on traffic data used non-parametric regressions for traffic flow forecasting (Zhong, J., Ling, S. 2015).

Another important part of analysis which needed to be conducted when performing data imputation is assessment of robustness (Zhong, M. et al. 2005). Zhong et al. found that modeles using additional observations as input and more sophisticated prediction techniques were able to make better imputations on a consistent basis.
